---
title: "Homework 1"
author: "Thomas Rodriguez"
format: html
embed-resources: true
editor: visual
---

## Question 1

I installed the three packages. But I did it in the console as I was instructed to do.

## Question 2

```{r}
library(causact)
library(dplyr)
library(igraph)

df <- dplyr::as_data_frame(x = c(1,2,3))

glimpse(df)

x <- c(5,6,2,7,9,1)
n_distinct(x)
```

The code that specifies using as_data_frame from dplyr works because the function from dplyr takes objects and turns them into a data frame, which is what we are doing with c(1, 2, 3).

R called the function from igraph instead of dplyr because it was loaded in after dplyr, and subsequently "masked" the as_data_frame object from dplyr, therefore becoming the primary reference point for that specific function when the package is unspecified.

## Question 3

```{r}
?n_distinct
```

The n_distinct function is used to count how many unique combinations are in a set of vectors. In reference to question 2, the function counted six, because in the value x, there are six different unique combinations: 5, 6, 2, 7, 9, and 1. If there were multiples of the same numbers, it would still be six because those numbers are not distinct.

## Question 4

```{r}
glimpse(baseballData)
```

The dataset has 12,145 rows and 5 columns. The Home column is a categorical variable while HomeScore is a quantitative variable.

## Question 5

```{r}
baseballData[1,]
baseballData[,2:3] %>% head()
```

One row in the data represents an observation of a baseball game.

These two columns are variables that represent the home team and the visitor team of each baseball game.

## Question 6

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)

df <- tibble( 
  Name = name, 
  Goals = goals, 
  Year_Started = year_started) 

glimpse(df)
```
