---
title: "Homework05"
author: "Thomas Rodriguez"
format: html
embed-resources: true
editor: visual
---

# Chapters 1-3 - Data Visualization

## Question 1

```{r}
#| message: false
library(tidyverse)
theme_set(theme_minimal())
# Read in the data 
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

I expect that people who record more exercise will have less BMI.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

There is a negative correlation between exercise and BMI.

```{r}
ggplot(exercise_data,
       aes(x = Exercise,
           y = BMI)) +
       geom_point()
```

The graph looks like a dinosaur.

## Question 2

```{r}
#| message: false
library(causact)
glimpse(corruptDF)
```

```{r}
?corruptDF
```

CPI2017 captures the Corruption Perceptions Index score for 2017 which demonstrates a country/territory's perceived public sector corruption on a scale of 0-100, where the higher the score the less corrupt. HDI2017 captures the Human Development Index for 2017 which measures achievement of a nation's longevity, education and income from a simple unweighted average.

## Question 3

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017)) +
  geom_point(alpha = .2)
```

The CPI and HDI for 2017 seem to have a positive relationship, meaning that country's that have a higher CPI score also have a higher HDI measure. However, they are not that strongly correlated as the data is relatively spread out.

## Question 4

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam")
```

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm")
```

The difference is that the gam method is not a straight line and follows the average of the points while the lm method is a straight line and demonstrates the steady increase. I prefer the lm method because it follows normal distribution more closely.

## Question 5

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017,
           color = region,
           fill = region)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm")
```

The graph is too cluttered to see the differences by region.

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017,
           color = region,
           fill = region)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  facet_wrap(~region)
```

## Question 6

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017,
           color = region,
           fill = region)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  facet_wrap(~region) +
  scale_x_reverse()
```

## Question 7

```{r}
ggplot(corruptDF,
       aes(x = HDI2017,
           y = CPI2017,
           color = region,
           fill = region)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  facet_wrap(~region) +
  scale_x_reverse() +
  labs(title = "Relationship between CPI and HDI 2017",
       subtitle = "by region",
       caption = "Data from corruptDF dataset in causact package")
```

## Question 8

```{r}
ggsave(filename = "CPI&HDIPlot.png")
```
